### 2019-12-31 main.tex
Update on Overleaf.

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{SBML L3 paper response to reviewers}
\author{Mike Hucka}
\date{December 2019}

\begin{document}

\maketitle

\section{Introduction}

\end{document}